Prediction is Hard (Especially About the Future)

### 2015-08-07 Manuscript.tex
Updates from @rahlk's local machine.

\usepackage{listings}
\usepackage{tabu}

\definecolor{lightgray}{gray}{0.8}
\definecolor{darkgray}{gray}{0.6}
\definecolor{Gray}{gray}{0.95}
\definecolor{LightGray}{gray}{0.975}

@@

Clustering is followed generating \textit{contrast sets}. These contrast sets represent recommendations on what could be altered to better improve an outcome. In this work we have explored several algorithms as possible tools to identify contrast between clusters. These fall into three broad categories: a) Case Based Reasoning techniques (Nearest Neighbors), b) A Gradient base planner (called HOW), and c) Decision Trees. These techniques are discussed below. It is worth noting that they are organized in such a way that each technique seeks to address certain fallacies in the ones that precede it.

@@

A classic example of a CBR is a nearest neighbor approach. The nearest neighbor approach is rather straight forward and has been developed as a "straw-man"; i.e., a simple baseline tool to act as a benchmark used to evaluate other methods.

Our approach is rather straight forward, we divide the $N$ training samples into $\sqrt{N}$ clusters and compute the centroid of these clusters. For every test case, we identify a cluster from the training set that most closely resembles it. Following this, we find the nearest cluster with a better performance score. The differences in the attributes between these two clusters constitute the \textit{"contrast set"}. These contrast sets acts as plans that can be used to reflect over the test cases to improve them.

@@

>>>>

\usepackage{listings}
\usepackage{tabu}
\DeclareMathOperator*{\argmin}{arg\,min} 
\DeclareMathOperator*{\argmax}{arg\,max}
\definecolor{lightgray}{gray}{0.8}
\definecolor{darkgray}{gray}{0.6}
\definecolor{Gray}{gray}{0.95}
\definecolor{LightGray}{gray}{0.975}

@@

Clustering is followed generating \textit{contrast sets}. These contrast sets represent recommendations on what could be altered to better improve an outcome. In this work we have explored several algorithms as possible tools to identify contrast between clusters. These fall into three broad categories: a) Case Based Reasoning techniques (Nearest Neighbors), b) A Gradient base planner (called HOW), and c) Decision Trees. These techniques are discussed below. It is worth noting that the following techniques are organized in such a way that each one seeks to address certain fallacies of the ones that precede it.

@@

A classic example of a CBR is K Nearest Neighbor. Our version of nearest neighbor for planning is rather straight forward. It has been developed as a "straw-man"; i.e., a simple baseline tool to act as a benchmark used to evaluate other methods.

We divide the $N$ training samples into $\sqrt{N}$ clusters and compute the centroid of these clusters. For every test case, we identify a cluster from the training set that most closely resembles it. Following this, we find the nearest cluster with a better performance score. The differences in the attributes between these two clusters constitute the \textit{"contrast set"}. These contrast sets acts as plans that can be used to reflect over the test cases to improve them.

@@

\begin{figure}[t!]
	\small
	~\hrule~

	{\bf \fig{where}a: top down clustering:}

	Data can be rapidly and recursively divided   as follows.
	Find   two   distance cases,  $X,Y$
	by picking any case $W$ at random, then setting $X$ to its most
	distant case, then setting $Y$ to the case most distant from
	$X$~\cite{fastmap}
	(this requires only $O(2N)$ comparisons
	of $N$ cases).
	Next, HOW projects each case $Z$
	onto a {\tt Slope} that  runs between $X,Y$ using the cosine
	rule of \fig{where}b. After that,  split the data at the median $x$ value of all cases and
	recurses on each half  (stopping when
	one half has less  than $\sqrt{N}$ of the original population).

	{\bf \fig{where}b: finding distances:}

	In the \fig{where}a, to compute distances, we use
	the Euclidean measure recommended for
	case-based reasoning by Aha et al.~\cite{aha91};
	i.e. $\sqrt{\sum_iw_i(X_i-Y_i)^2}$ where $X_i,Y_i$
	where values are  normalized 0..1 for the range min..max and 
	$w_i$ is some importance weight given to attribute $i$.
	Usually, $w_i=1$ but it can be adjusted using, say,
	the feature weighting methods of \fig{where}d.

	{\bf \fig{where}c: 3 point geometry:}

	Let   $X,Y$ be two points joined by  a {\em Slope} of  length $c$.
	A third point, $Z$, has distances  $a=dist(Z,X)$ and
	$b=dist(Z,Y)$ to $X,Y$, respectively.
	According to the cosine rule,   $Z$ projects onto  $\overline{XY}$
	at distance $x=(a^2 + c^2 - b^2)/(2c)$ from $X$.
	Further, according to Pythagoras' theorem, $Z$ stands at a distance
	$y = \sqrt{a^2 - x^2}$ from the line $\overline{XY}$.

	{\bf \fig{where}d: feature weighting:}

	HERE's feature weighting algorithm
	comes from the CART regression tree learner~\cite{Breiman1984}.
	It sorts independent variables
	according to how well they can reduce the variability
	of a  numeric objective.
	If we split some column $i$ of independent numeric data  into $N=n_1 + n_2 + ..$ splits,
	then the expected
	value of the objective value of the split  is $w_i = \sum_i v_in_i/N$
	where $v$ is the standard deviation of the objective.
	A recursive procedure can  find those divisions $n_1,n_2,...$ by (a)~sorting a numeric column,
	then split at $\argmin_i w_i$, then recursing into each half.

	\caption{Sub-routines within HOW.}\label{fig:where}
\end{figure}
